Hydras \& Co.

### 2023-07-11 doc/chapter-fol.tex
use latex package turnstile
work on chapter-natural-deduction

@@

@@

\inputsnippets{FolExamples/PrfEx2}

@@

\inputsnippets{FolExamples/PrfEx3}

@@

Indeed, in the next chapter, we will consider a proof system
\texttt{SysPrf}, based on \texttt{Prf}, the properties of which will 
allow us to prove theorems in a very simpler way.

>>>>

We will also use (in the text) the notation $l \hdash{} F$ for the type \texttt{Prf $l$ $F$}.

@@

\subsubsection{Notation}

@@

\label{proof:PrfEx2}
\inputsnippets{FolExamples/PrfEx2}

@@

\label{proof:PrfEx3}
\inputsnippets{FolExamples/PrfEx3}

@@

Indeed, in the next chapter, we will consider a proof system
\texttt{SysPrf}, based on \texttt{Prf}, the properties of which will 
allow us to prove theorems in a much simpler way.

### 2023-07-11 doc/chapter-natural-deduction.tex
work on chapter-natural-deduction

The answer to both questions in~\cite{Goedel}  is the definition of a proof system, derived from \texttt{Prf}, which considers
\emph{sets of hypotheses} (called \emph{systems} in ~\cite{Goedel}) instead of
\emph{list of hypotheses}, thus making abstraction of
repetition and order of appearance of hypotheses in the context.

The new system is simply defined as below:
\inputsnippets{folProof/SysPrfDef}

@@

\begin{itemize}
\item  $\Gamma \sdash A$ for
  (\texttt{SysPrf $L$ $\Gamma$ $A$}).
\item $\Gamma, \Delta$ for $\Gamma\cup\Delta$
\item $\Gamma, A,\dots B$ for $\Gamma\cup\{A \dots B\}$
\end{itemize}

@@

In the following script, we use \texttt{PrfEx2}
as a witness for proving a set-based version of the 
original proof term.

$$A \sststile{\mathrm{S}}{} A$$

@@

Prove the following lemma:
\inputsnippets{FolExamples/MPSys}

@@

\section{The deduction theorem}

@@

@@

\subsection{Example}

For instance, let us prove Peirce's rule.
 (in  ~\href{../theories/html/hydras.MoreAck.FolExamples.html}{MoreAck.FolExamples} ).

\subsubsection{Prelude}

%\inputsnippets{folPeirce/prelude}

\subsection{Proof of Peirce's law}

@@

Let us try to do an implication introduction.

@@

Now, we may use the law of excluded middle with the formula $P$. The only non-trivial case is about $\sim P$.

>>>>

The answer to both questions in~\cite{Goedel}  is the definition of a proof system, derived from \texttt{Prf}, which considers
\emph{sets of hypotheses} (called \emph{systems} in ~\cite{Goedel}) instead of
\emph{list of hypotheses}, thus making abstraction of
the repetition and order of appearance of hypotheses in the context.

The new system is simply defined as below (for  a given language $L$)\,\footnote{In some shown snippets, arguments like $L$ may be or not be implicit (depending on the section they are extracted from). Please look at the \coq source.}.

\inputsnippets{folProof/SysPrfDef}

@@

In the text, we may use the abbreviation $\Gamma \sdash A$ for
(\texttt{SysPrf $L$ $\Gamma$ $A$}) and
$\sdash A$ for
(\texttt{SysPrf $L$ Empty\_set $A$})

We may also omit obvious braces in some set expressions:

\begin{itemize}

\item $\Gamma, \Delta$ for $\Gamma\cup\Delta$,
\item $A, B, C$ for $\{A, B, C\}$,
\item $\Gamma, A,\dots B$ for $\Gamma\cup\{A \dots B\}$
\item \emph{etc.}
\end{itemize}

@@

In the following script, we use \texttt{PrfEx2}
as a witness for proving a set-based version of the 
original proof term, namely $A; A \arrow B; A \arrow B \arrow C \sdash C$.

@@

Prove the following lemma (without the handy lemmas from
\href{../theories/html/hydras.Ackermann.folLogic.html}{Ackermann.folLogic} and their corollaries!).

\inputsnippets{FolExamples/MPSys}

@@

\section{The Deduction theorem}

The deduction theorem (proved in \href{../theories/html/hydras.Ackermann.Deduction.html}{Ackermann.Deduction} is a handy 
tool for proving a proposition $A\arrow B$ by pushing 
the hypothesis $A$ into the context (it corresponds to 
implication introduction in \coq).

@@

\subsection{sketch of proof}

We advise the reader to replay this proof on h.er.is computer
 in order to better understand its structure.

Let us give only a few comments.

Let us assume the hypothesis \texttt{H: $T, g \sdash f$},
meaning that there exists some list $F$ whose elements 
belong to $T \cup \{g\}$, and a proof-tree $t$ of type
$F \vdash h$.

The heart of the proof is an induction on $t$ proving 
$F \cap T \sdash (g\arrow f)$ \footnote{Please forgive the implicit coercion from lists to sets!}.

Please note that the case $S=\texttt{nil}$ is common to 
many constructors of \texttt{Prf}, thus the proof script starts
with a study of this particular case, simply applied 11 times in the rest of the proof.

@@

\subsection{Example: proof of Peirce's law}

For instance, let us prove Peirce's rule, \emph{i.e.}
$\sdash ((A \arrow B) \arrow A) \arrow A$ for any formulas $A$ and $B$.

 The proof in \coq is available in  ~\href{../theories/html/hydras.MoreAck.FolExamples.html}{MoreAck.FolExamples}.

@@

Let us start with  an implication introduction.

@@

Now, we may eliminate the disjunction 
in the instance $\sim A \vee A$ of the law of excluded middle.
The only non-trivial case is about $\sim A$.
